---
title: 'Reproducible Research: Peer Assessment 1'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---


## Loading and preprocessing the data
```{r, echo = TRUE}
activity<-read.csv(file = "activity.csv", header = T)

```


## What is mean total number of steps taken per day?
```{r}
library(dplyr)
totStepsPerDay<-
  activity %>%
  group_by(date) %>%
  summarize(totSteps=sum(steps,na.rm = T))
plot(x=totStepsPerDay$date,y=totStepsPerDay$totSteps, xlab="Date", ylab="Number of steps")
meanSteps<-mean(totStepsPerDay$totSteps)
medSteps<-median(totStepsPerDay$totSteps)
```
The mean of the steps is `r meanSteps` and the median is `r medSteps`.



## What is the average daily activity pattern?
```{r}
meanStepsPerInterval<-
  activity %>%
  group_by(interval) %>%
  summarize(meanSteps=mean(steps,na.rm = T))
plot(x = meanStepsPerInterval$interval,y=meanStepsPerInterval$meanSteps, type = "l", xlab = "Interval", ylab = "Mean steps")
maxStepsInterval<-meanStepsPerInterval[which.max(meanStepsPerInterval$meanSteps),1]
```
The interval that contains the maximum number of steps is `r maxStepsInterval`.

## Imputing missing values
```{r}
numNA<-sum(!complete.cases(activity))
```
The missing values are `r numNA`.
```{r}
#find the index of the NA's
indexIncomplete<-which(!complete.cases(activity))
#create new dataframe with index and the corresponding interval
dfReplace<-data.frame(indexIncomplete,interval=activity$interval[indexIncomplete])
#add a column to the dataframe merging with the mean for the corresponding interval
dfReplace<-merge(dfReplace,meanStepsPerInterval,by.x = "interval", by.y = "interval" )
#sort the dataframe by index
dfReplace<-dfReplace[order(dfReplace$indexIncomplete),]
#make new complete dataframe based on the incomplete one
activityComplete<-activity
#assign the value for the completed dataframe
activityComplete$steps[indexIncomplete]<-dfReplace$meanSteps

totStepsPerDayComplete<-
  activityComplete %>%
  group_by(date) %>%
  summarize(totSteps=sum(steps,na.rm = T))
plot(x=totStepsPerDayComplete$date,y=totStepsPerDayComplete$totSteps, xlab="Date", ylab="Number of steps")
meanStepsComplete<-mean(totStepsPerDayComplete$totSteps)
medStepsComplete<-median(totStepsPerDayComplete$totSteps)
```
The mean of the steps were `r meanSteps` and now is `r meanStepsComplete`. The median were `r medSteps` and now is `r medStepsComplete`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekday<-weekdays(as.Date(activityComplete$date))
weekday[weekday %in% c("lunedì","martedì","mercoledì","giovedì","venerdì")]<-"weekday"
weekday[weekday %in% c("sabato","domenica")]<-"weekend"
activityComplete$weekday<-weekday
activityComplete$weekday<-as.factor(activityComplete$weekday)

meanStepsPerIntervalComplete<-
  activityComplete %>%
  group_by(interval,weekday) %>%
  summarize(meanSteps=mean(steps,na.rm = T))

xyplot(meanSteps~interval | weekday ,data=meanStepsPerIntervalComplete)
```
